---
title: "Assignment Template"
format: html
---

# Question 1

```{r}
# install.packages("invgamma")
# install.packages("extraDistr")

library(invgamma)
library(mvtnorm)
library(ggplot2)
library(MASS)
library(extraDistr)
```

```{r}
# Packages required
require(MASS)

# Load the Auto MPG dataset
auto <- read.csv("auto.csv")

auto <- na.omit(auto)

Y <- matrix(auto$mpg, ncol = 1)

Xpreds <- auto[, c("displacement", "horsepower", "weight", "acceleration")]


X <- as.matrix(cbind(1, scale(Xpreds)))

n <- nrow(X)

```

## Gibbs Sampling

## Helper functions

```{r}
b_tilde = matrix(0,nrow = 5,1)

M = diag(1,5)

mu_b_function = function(M, X, b_hat, b_tilde){
  solve((M + t(X)%*%X)) %*% (t(X)%*%X%*%b_hat)
}


A2_function = function(y,b_tilde,mu_b){
  (t(y)%*%y)  - (t(mu_b)%*%(M+t(X)%*%X)%*%mu_b)
}

b_hat_function = function(X,Y){
  solve(t(X)%*%X)%*%t(X)%*%Y
}


b_hat = b_hat_function(X,Y)

mu_b = mu_b_function(M,X,b_hat,b_tilde)

A2 = A2_function(Y,b_tilde,mu_b)
```

### Conditional Posteriors

```{r}
beta_posterior = function(mu_b,sigma){
  var = sigma*solve((M+t(X)%*%X))
}

sigma_posterior = function(a,b,A2) {
  
}
```

### Sampling from Inverse Gamma

-   Use to get $\sigma^2$

```{r}
#| eval: false
#| include: false
#Fixed for all iterations
a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)


sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
```

### Sample from Multivariate Normal

-   To get $\beta$

```{r}
#| eval: false
#| include: false
sigma = sigma_hat*solve((M+t(X)%*%X))
sigma
mean = mu_b
mean
beta_hat <- rmvnorm(n = 1, mean , sigma)
beta_hat

sample = c(sigma_hat,beta_hat)
names(sample) = c( "sigma", "b0", "b1", "b2", "b3", "b4")
sample
```

## Gibbs Sampling Function

```{r}
gibs_sampler = function(iterlim,A2,mu_b,M){
  
  #Create matrix to store values 
  sample_matrix = matrix(0,nrow = iterlim,ncol = 6)
  colnames(sample_matrix) = c("sigma", "b0", "b1", "b2", "b3", "b4")
  
  #Initialise values of sigma^2 and beta 
  a = 1
  b = 1
  shape = a + (n/2)
  rate = b + (A2/2)
  sigma_hat <- rinvgamma(1, shape, rate)
  
  sigma = sigma_hat*solve((M+t(X)%*%X))
  mean = mu_b
  beta_hat = rmvnorm(n = 1, mean , sigma)
  
  #Store intial values
  sample = c(sigma_hat,beta_hat)
  
  sample_matrix[1,] = sample
  
  
  
  for(i in 2:iterlim){
    
    #Generate value for sigma^2
    sigma_hat <- rinvgamma(1, shape, rate)
    
    #Generate value for beta 
    sigma = sigma_hat*solve((M+t(X)%*%X))
    mean = mu_b
    beta_hat = rmvnorm(n = 1, mean , sigma)
    
    #Store sample
    sample = c(sigma_hat,beta_hat)
  
    sample_matrix[i,] = sample
    
    
  }
  
  return(sample_matrix)
    
}
```

## Generate Sample

```{r}
mat = gibs_sampler(50000,A2,mu_b,M)
```

## Trace plots 

### $\sigma^2$

```{r}
x = seq(1,50000)
sigma_squared = mat[,1]

pdf("trace_sigma.pdf")
plot(x,sigma_squared,type = "l", main = expression(paste("Trace Plot for" ~ sigma^2)),
     ylab = expression(sigma^2), 
     xlab = "Iteration")
#dev.off()
```

### $\beta_0$

```{r}
b0 = mat[,2]
pdf("trace_b0.pdf")
plot(x,b0,type = "l", main = expression(paste("Trace Plot for" ~ beta[0])),
     ylab = expression(beta[0]), 
     xlab = "Iteration")
```

### $\beta_1$

```{r}
b1 = mat[,3]
pdf("trace_b1.pdf")
plot(x,b1,type = "l", main = expression(paste("Trace Plot for" ~ beta[1])),
     ylab = expression(beta[1]), 
     xlab = "Iteration")
```

### $\beta_2$

```{r}
b2 = mat[,4]
pdf("trace_b2.pdf")
plot(x,b2,type = "l", main = expression(paste("Trace Plot for" ~ beta[2])),
     ylab = expression(beta[2]), 
     xlab = "Iteration")
```

### $\beta_3$

```{r}
b3 = mat[,5]
pdf("trace_b3.pdf")
plot(x,b3,type = "l", main = expression(paste("Trace Plot for" ~ beta[3])),
     ylab = expression(beta[3]), 
     xlab = "Iteration")
```

### $\beta_4$

```{r}
b4 = mat[,6]
pdf("trace_b4.pdf")
plot(x,b4,type = "l", main = expression(paste("Trace Plot for" ~ beta[4])),
     ylab = expression(beta[4]), 
     xlab = "Iteration")
```

## Density Plots

### $\sigma^2$

```{r}
hist(sigma_squared, probability = TRUE,breaks = 50, 
     main = expression(paste("Posterior of" ~ sigma^2)) , xlab = expression(sigma^2))

a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)

#Overlay Inverse Gamma
curve(dinvgamma(x, shape,rate),
      from = 14, to = 24,
      add = TRUE, col = "red", lwd = 2)

abline(v = mean(sigma_squared), col = "blue", lwd = 2, lty = 2)
```

### $\beta_0$

```{r}
pdf("hist_b0.pdf")
hist(b0,breaks = 50, main = expression(paste("Density plot of" ~ beta[0[]])) , xlab = expression(beta[0]))

#Posterior mean
abline(v = mean(b0), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[1], col = "red", lwd = 2, lty = 2)

```

### $\beta_1$

```{r}
pdf("hist_b1.pdf")
hist(b1,breaks = 50, main = expression(paste("Density plot of" ~ beta[1])) , xlab = expression(beta[1]))

#Posterior mean
abline(v = mean(b1), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[2], col = "red", lwd = 2, lty = 2)
```

### $\beta_2$

```{r}
pdf("hist_b2.pdf")
hist(b2,breaks = 50, main = expression(paste("Density plot of" ~ beta[2])) , xlab = expression(beta[2]))

#Posterior mean
abline(v = mean(b2), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[3], col = "red", lwd = 2, lty = 2)
```

### $\beta_3$

```{r}
pdf("hist_b3.pdf")
hist(b3,breaks = 50, main = expression(paste("Density plot of" ~ beta[3])) , xlab = expression(beta[3]))

#Posterior mean
abline(v = mean(b3), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[4], col = "red", lwd = 2, lty = 2)
```

### $\beta_4$

```{r}
pdf("hist_b4.pdf")
hist(b4,breaks = 50, main = expression(paste("Density plot of" ~ beta[4])) , xlab = expression(beta[4]))

#Posterior mean
abline(v = mean(b4), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[5], col = "red", lwd = 2, lty = 2)
```

## Credibility Intervals

### $\sigma^2$

```{r}
cred_ints = quantile(mat[, "sigma"], probs = c(0.025, 0.975))
```

### $\beta$

```{r}
cred_int_B = apply(mat[, c("b0","b1","b2","b3","b4")], 2, quantile, probs = c(0.025, 0.975))
```

## Confidence Intervals

-   Fit linear model and compute confidence intervals

```{r}
fit = lm(Y~X)

summary(fit)

confints = confint(fit)
```

## Density Plots w/ mean, OLS, Cred Intervals and CI

### Overlayed Density plot of $\sigma^2$

```{r}
hist(sigma_squared, probability = TRUE,breaks = 50, 
     main = expression(paste("Posterior of" ~ sigma^2)) , xlab = expression(sigma^2))

a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)

#Overlay Inverse Gamma
curve(dinvgamma(x, shape,rate),
      from = 14, to = 40,
      add = TRUE, col = "red", lwd = 2)

abline(v = mean(sigma_squared), col = "blue", lwd = 2, lty = 2)
abline(v = cred_ints, col = "darkmagenta", lwd =2, lty = 2)
```

### Overlayed Density Plot for $\beta_0$

```{r}
pdf("hist_b0_overlay.pdf")
hist(b0,breaks = 50, main = expression(paste("Density plot of" ~ beta[0[]])) , xlab = expression(beta[0]))

#Posterior mean
abline(v = mean(b0), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[1], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 1], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[1, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_1$

```{r}
pdf("hist_b1_overlay.pdf")
hist(b1,breaks = 50, main = expression(paste("Density plot of" ~ beta[1])) , xlab = expression(beta[1]))

#Posterior mean
abline(v = mean(b1), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[2], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 2], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[3, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_2$

```{r}
pdf("hist_b2_overlay.pdf")
hist(b2,breaks = 50, main = expression(paste("Density plot of" ~ beta[2])) , xlab = expression(beta[2]))

#Posterior mean
abline(v = mean(b2), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[3], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 3], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[4, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_3$

```{r}
pdf("hist_b3_overlay.pdf")
hist(b3,breaks = 50, main = expression(paste("Density plot of" ~ beta[3])) , xlab = expression(beta[3]))

#Posterior mean
abline(v = mean(b3), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[4], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 4], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[5, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_4$

```{r}
pdf("hist_b4_overlay.pdf")
hist(b4,breaks = 50, main = expression(paste("Density plot of" ~ beta[4])) , xlab = expression(beta[4]))

#Posterior mean
abline(v = mean(b4), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[5], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 5], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[6, ], col = "darkmagenta", lwd = 2, lty = 2)
```

## Informative Prior

```{r}
b_tilde = matrix(c(23,-3,-5,-5,0),nrow = 5,1)


M = 10*diag(1,5)



mu_b = mu_b_function(M,X,b_hat,b_tilde)

A2 = A2_function(Y,b_tilde,mu_b)

```

### Sample

```{r}
mat2 = gibs_sampler(50000,A2,mu_b,M)

#Extract vectors 
sigma_squared = mat2[,1]
b0 = mat2[,2]
b1 = mat2[,3]
b2 = mat2[,4]
b3 = mat2[,5]
b4 = mat2[,6]
```

### Density Plots

### $\sigma^2$

```{r}
hist(sigma_squared, probability = TRUE,breaks = 50, 
     main = "Posterior of sigma^2", xlab = "sigma^2")

a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)

#Overlay Inverse Gamma
curve(dinvgamma(x, shape,rate),
      from = 25, to = 40,
      add = TRUE, col = "red", lwd = 2)



abline(v = mean(sigma_squared), col = "blue", lwd = 2, lty = 2)
```

### $\beta_0$

```{r}
hist(b0,breaks = 50)

#Posterior mean
abline(v = mean(b0), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[1], col = "red", lwd = 2, lty = 2)

```

### $\beta_1$

```{r}
hist(b1,breaks = 50)

#Posterior mean
abline(v = mean(b1), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[2], col = "red", lwd = 2, lty = 2)
```

### $\beta_2$

```{r}
hist(b2,breaks = 50)

#Posterior mean
abline(v = mean(b2), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[3], col = "red", lwd = 2, lty = 2)
```

### $\beta_3$

```{r}
hist(b3,breaks = 50)

#Posterior mean
abline(v = mean(b3), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[4], col = "red", lwd = 2, lty = 2)
```

### $\beta_4$

```{r}
hist(b4,breaks = 50)

#Posterior mean
abline(v = mean(b4), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[5], col = "red", lwd = 2, lty = 2)
```

# Question 2

## Data Generating Function

```{r}
#### Data Generating Functions ####


generate_lost <- function(grid_size, nsims){
  
  # Function to generate the prior distribution for the 
  # location of the lost fisherman
  # Args: 
  #       grid_size: the dimensions of the square search grid
  #       nsims: number of samples to base the prior distribution on
  
  mu_vec  <- c(grid_size/2, grid_size/2)
  sig_mat <- matrix(c(2, 1, 5, 5), 2,2)
  
  dat <- mvrnorm(nsims, mu_vec, sig_mat)
  dat <- round(abs(dat))
  
  prior <- matrix(rep(0,grid_size^2), grid_size, grid_size)
  for (i in 1:NROW(dat)){
    
    if (dat[i,1] < grid_size & dat[i,2] < grid_size){
      prior[dat[i,1], dat[i,2]] <- prior[dat[i,1], dat[i,2]] + 1
    }
    
  }
  prior <- prior/sum(prior)
  return(prior)
}


generate_fisherman <- function(grid_size){
  
  # Function to generate the true location of the lost fisherman.
  # This should not effect the search decision in any way!! It is unkown
  # to the search crew.
  # Args: 
  #       grid_size: the dimensions of the square search grid

  
  mu_vec  <- c(grid_size/2, grid_size/2)
  sig_mat <- matrix(c(2, 1, 5, 5), 2,2)
  
  location  <- round(mvrnorm(1, mu_vec, sig_mat))
  true_grid <- matrix(rep(0, grid_size^2), grid_size, grid_size)
  true_grid[location[1], location[2]] <- 1
  
  return(true_grid)
}

#### Simulation ####

search_size <- 20
unifs <- runif(search_size^2, min = 0.6, max = 0.9)

#pi
detect_pr  <- matrix(unifs, ncol = search_size)



```

## Simulation

### Generate fisherman's true location

```{r}
#Store in here
true_loc = generate_fisherman(20)

true_ind <- which(true_loc == 1, arr.ind = TRUE)
```

### Gain prior distribution for fisherman's location using expert

```{r}
expert_prior = generate_lost(20,100)
```

### Search process

-   Search cell with highest probability of successful detection

### Loop

```{r}

result = 0

numiterations = 0

stages = c(1,5,10,15)

list_stages = list()

#Posterior probability 
true_cell_probs = c()

merged_dists = c()

while(result == 0){
  #Update the number of iterations
  numiterations = numiterations + 1
  
  if (numiterations %in% stages)  list_stages[[as.character(numiterations)]] = expert_prior
  
  #Merge distributions 
  merged_dists = expert_prior*detect_pr

  #Search cell with highest probability of occurence
  
  #Coordinates of that cell 
  highest_prob_coord = arrayInd(which.max(merged_dists),dim(merged_dists))

  #Detection probability
  pi = detect_pr[highest_prob_coord]
  
  #True state
  yi = true_loc[highest_prob_coord]

  
  #Simulate search process using bernoulli distribution
  #Zi|Yi
  result = rbinom(1,1,yi*pi)
  
  #Update occurence probabilities if fishermen not found
  
  #Will technically exit loop as soon as result = 1
  if(result == 0){
  old_theta = expert_prior[highest_prob_coord]
  
      
  # Update all other cells first (equation 3)
  expert_prior = expert_prior / (1 - pi * old_theta)
  
    
  #Update current cell (equation 2)
  expert_prior[highest_prob_coord] = (1 - pi) * old_theta / (1 - pi * old_theta)
  }
  
  #Posterior probability of occurence after exploring a specific 
  true_cell_probs <- c(true_cell_probs, expert_prior[true_ind])
    
  
}

list_stages[["Final"]] <- expert_prior

cat("It took",numiterations,"hours to find the fishermen")
```

## Heatmaps of occurrence probability

```{r}
list_stages[["Final"]] <- expert_prior

for (nm in names(list_stages)) {
  mat <- list_stages[[nm]]
  image(mat,
        main = paste("Iteration:", nm),
        col  = hcl.colors(64, "YlOrRd", rev = TRUE),
        zlim = c(0, max(sapply(list_stages, max))))
  grid(nx = ncol(mat), ny = nrow(mat), col = "black", lty = 1, lwd = 0.5)
  points(
    (true_ind[1] - 0.5) / nrow(mat),
    (true_ind[2] - 0.5) / ncol(mat),
    pch = 4, cex = 2, col = "blue", lwd = 2
  )
  
}
```

## Posterior Probability of Occurrence

```{r}
steps = seq(1,numiterations)

library(ggplot2)

df <- data.frame(step = steps, probability = true_cell_probs)

pdf("posterior_prob.pdf")
ggplot(df, aes(x = step, y = probability)) +
  geom_line(color = "#2196F3", linewidth = 1) +
  labs(title = "Posterior Probability of Occurence in True Location",
       x = "Iteration", y = "Probability") +
  theme_minimal()
```

## Comparison to Naive Strategy

### True fishermen location for both

```{r}
true_loc = generate_fisherman(20)
```

### Naive

```{r}
rows = seq(1,20)
columns = seq(1,20)

cells = expand.grid(rows, columns)

# Randomize grid
cells = cells[sample(nrow(cells)), ]

result = 0
numiterations_random = 1

while(result == 0){

  row    = cells[numiterations_random,1]
  column = cells[numiterations_random,2]

  result = true_loc[row,column]

  numiterations_random = numiterations_random + 1
}

cat("It took",numiterations_random - 1,"hours to find the fishermen under the naive strategy")

```

### Bayes

```{r}

search_size <- 20

unifs <- runif(search_size^2, min = 0.6, max = 0.9)

#pi
detect_pr  <- matrix(unifs, ncol = search_size)

expert_prior = generate_lost(20,100)

result = 0

numiterations = 0

merged_dists = c()

while(result == 0){
  #Update the number of iterations
  numiterations = numiterations + 1

  
  #Merge distributions 
  merged_dists = expert_prior*detect_pr

  #Search cell with highest probability of occurence
  
  #Coordinates of that cell 
  highest_prob_coord = arrayInd(which.max(merged_dists),dim(merged_dists))

  #Detection probability
  pi = detect_pr[highest_prob_coord]
  
  #True state
  yi = true_loc[highest_prob_coord]

  
  #Simulate search process using bernoulli distribution
  #Zi|Yi
  result = rbinom(1,1,yi*pi)
  
  #Update occurence probabilities if fishermen not found
  
  #Will technically exit loop as soon as result = 1
  if(result == 0){
  old_theta = expert_prior[highest_prob_coord]
  
      
  # Update all other cells first (equation 3)
  expert_prior = expert_prior / (1 - pi * old_theta)
  
    
  #Update current cell (equation 2)
  expert_prior[highest_prob_coord] = (1 - pi) * old_theta / (1 - pi * old_theta)
  }
  
    
  
}
cat("It took",numiterations,"hours to find the fishermen under the Bayes strategy")
```

# Question 3

### Preamble

```{r}
# Use the same Auto MPG data from Q1
# Y, X, X.c, n, k should already be in your environment from Q1.
# If not , re -run the Q1 preamble code .
# Contaminate the data with outliers
set.seed (12)
n1 <- 350 # Number of 'standard ' observations
n2 <- n - n1 # Number of outliers
outlier.idx <- ( n1 + 1) : n # Index set I2
# Inflate the response for the outlier observations
Y.contaminated <- Y
Y.contaminated[outlier.idx] <- Y[outlier.idx] +
rnorm(n2,mean = 0 , sd = 15)

```

### Separating data

```{r}
X_1 = X[1:n1,]
X_2 = X[outlier.idx,]

Y_1 = Y.contaminated[-outlier.idx]
Y_2 = Y.contaminated[outlier.idx]
```

## $\pi(\beta| \tau_1,\tau_2,X,y)$

-   Function to sample from conditional distribution of $\beta$

```{r}
T0 = diag(1,5)
sample_beta = function(tau_1,tau_2) {
  
  #Covariance matrix
  sigma_b = ((tau_1 * (t(X_1)%*%X_1 ))+ (tau_2 * (t(X_2) %*% X_2)) + solve(T0))
  
  sigma_b_inv = solve(sigma_b)
  
  #Mean vector
  mu_b = sigma_b_inv%*%((tau_1 * t(X_1)%*%Y_1) + (tau_2*t(X_2)%*%Y_2))
  
  #Sample and return
  beta = rmvnorm(n = 1, mu_b , sigma_b_inv)
  beta = matrix(beta,5)
  return(beta)
  
}
```

## $\pi(\tau_1|\beta,\tau_2,X,y)$

-   Function to sample from conditional distribution of $\tau_1$

```{r}
sample_tau_1 = function(beta){
  
  #Shape Parameter
  a = 1
  a1 = (n1/2 + a)

  #Scale Parameter
  S1 = 0.5 * ((t(Y_1)%*%Y_1)-(2*t(beta)%*%t(X_1)%*%Y_1)+(t(beta)%*%t(X_1)%*%X_1%*%beta))
  
  b = 1
  b1 = b + S1

  tau_1 = rgamma(1,shape = a1,scale = 1/b1)
  
  return(tau_1)
  
}

```

## $\pi(\tau_2|\beta,\tau_1,X,y)$

-   Function to sample from conditional distribution of $\tau_2$

```{r}
sample_tau_2 = function(beta){
  
  #Shape Parameter
  a = 1
  a2 = (n2/2 + a)

  #Scale Parameter
  S2 = 0.5 * ((t(Y_2)%*%Y_2)-(2*t(beta)%*%t(X_2)%*%Y_2)+(t(beta)%*%t(X_2)%*%X_2%*%beta))
  
  b = 1
  b2 = b + S2

  tau_2 = rgamma(1,shape = a2,scale = 1/b2)
  
  return(tau_2)
  
}
```

## Gibbs Sampling Function

```{r}
gibbs_sampler_2 = function(iterlim) {
    #Create matrix to store values 
  sample_matrix = matrix(0,nrow = iterlim,ncol = 7)
  colnames(sample_matrix) = c("tau_1","tau_2", "b0", "b1", "b2", "b3", "b4")
  
  #Initialise values of beta,tau_1 and tau_2 
  tau_2 = runif(1,0,1)
  
  tau_1 = tau_2 + runif(1,0,1)
  
  beta = matrix(runif(5,-5,5),5)
  
  #Store intial values
  sample = c(tau_1,tau_2,beta)
  
  sample_matrix[1,] = sample
  
  #Loop and sample values
  for(i in 2:iterlim){
    
    #Generate value for tau_1
    tau_1 = sample_tau_1(beta)
    
    #Generate value for tau_2
    tau_2 = sample_tau_2(beta)
    
    #Generate value for beta 
    beta = sample_beta(tau_1,tau_2)
    
    #Store sample
    sample = c(tau_1,tau_2,beta)
  
    sample_matrix[i,] = sample
    
    
  }
  
  return(sample_matrix)
  
  
}
```

## Generate Sample

```{r}
mat3 = gibbs_sampler_2(150000)

burn_in_index = seq(1,100000)
mat3 = mat3[-burn_in_index,]
```

### $\beta_0$

```{r}
x = seq(1, 50000)
b0 = mat3[,3]
pdf("post_trace_b0.pdf")
plot(x, b0, type = "l",
     main = expression(paste("Trace Plot for" ~ beta[0])),
     ylab = expression(beta[0]),
     xlab = "Iteration",
     ylim = range(b0))

```

### $\beta_1$

```{r}
b1 = mat3[,4]
pdf("post_trace_b1.pdf")
plot(x,b1,type = "l", main = expression(paste("Trace Plot for" ~ beta[1])),
     ylab = expression(beta[1]), 
     xlab = "Iteration")
```

### $\beta_2$

```{r}
b2 = mat3[,5]
pdf("post_trace_b2.pdf")
plot(x,b2,type = "l", main = expression(paste("Trace Plot for" ~ beta[2])),
     ylab = expression(beta[2]), 
     xlab = "Iteration")
```

### $\beta_3$

```{r}
b3 = mat3[,6]
pdf("post_trace_b3.pdf")
plot(x,b3,type = "l", main = expression(paste("Trace Plot for" ~ beta[3])),
     ylab = expression(beta[3]), 
     xlab = "Iteration")
```

### $\beta_4$

```{r}
b4 = mat3[,7]
pdf("post_trace_b4.pdf")
plot(x,b4,type = "l", main = expression(paste("Trace Plot for" ~ beta[4])),
     ylab = expression(beta[4]), 
     xlab = "Iteration")
```

### $\tau_1$

```{r}
t1 = mat3[,1]
pdf("post_trace_t1.pdf")
plot(x,t1,type = "l", main = expression(paste("Trace Plot for" ~ tau[1])),
     ylab = expression(tau[1]), 
     xlab = "Iteration")
```

### $\tau_2$

```{r}
t2 = mat3[,2]
pdf("post_trace_t2.pdf")
plot(x,t1,type = "l", main = expression(paste("Trace Plot for" ~ tau[2])),
     ylab = expression(tau[2]), 
     xlab = "Iteration")
```

### Density Plots

### $\beta_0$

```{r}
hist(b0,breaks = 50, main = expression(paste("Density plot of" ~ beta[0[]])) , xlab = expression(beta[0]))

#Posterior mean
abline(v = mean(b0), col = "blue", lwd = 2, lty = 2)

```

### $\beta_1$

```{r}
hist(b1,breaks = 50, main = expression(paste("Density plot of" ~ beta[1])) , xlab = expression(beta[1]))

#Posterior mean
abline(v = mean(b1), col = "blue", lwd = 2, lty = 2)
```

### $\beta_2$

```{r}
hist(b2,breaks = 50, main = expression(paste("Density plot of" ~ beta[2])) , xlab = expression(beta[2]))

#Posterior mean
abline(v = mean(b2), col = "blue", lwd = 2, lty = 2)

```

### $\beta_3$

```{r}
hist(b3,breaks = 50, main = expression(paste("Density plot of" ~ beta[3])) , xlab = expression(beta[3]))

#Posterior mean
abline(v = mean(b3), col = "blue", lwd = 2, lty = 2)

```

### $\beta_4$

```{r}
hist(b4,breaks = 50, main = expression(paste("Density plot of" ~ beta[4])) , xlab = expression(beta[4]))

#Posterior mean
abline(v = mean(b4), col = "blue", lwd = 2, lty = 2)
```
