---
title: "Assignment Template"
format: html
---

```{r}
#install.packages("invgamma")

library(invgamma)
library(mvtnorm)
```

```{r}
# Packages required
require(MASS)

# Load the Auto MPG dataset
auto <- read.csv(
  "https://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data",
  header = FALSE,
  sep = "",   # whitespace-separated file
  col.names = c("mpg", "cylinders", "displacement", "horsepower",
                "weight", "acceleration", "model_year", "origin", "car_name"),
  na.strings = "?"
)

auto <- na.omit(auto)

Y <- matrix(auto$mpg, ncol = 1)

Xpreds <- auto[, c("displacement", "horsepower", "weight", "acceleration")]

X <- as.matrix(cbind(1, scale(Xpreds)))

n <- nrow(X)

```

```{r}
#### Required Functions ####

library(ggplot2)
library(MASS)

#### Data Generating Functions ####

generate_lost <- function(grid_size, nsims){
  
  # Function to generate the prior distribution for the 
  # location of the lost fisherman
  # Args: 
  #       grid_size: the dimensions of the square search grid
  #       nsims: number of samples to base the prior distribution on
  
  mu_vec  <- c(grid_size/2, grid_size/2)
  sig_mat <- matrix(c(2, 1, 5, 5), 2,2)
  
  dat <- mvrnorm(nsims, mu_vec, sig_mat)
  dat <- round(abs(dat))
  
  prior <- matrix(rep(0,grid_size^2), grid_size, grid_size)
  for (i in 1:NROW(dat)){
    
    if (dat[i,1] < grid_size & dat[i,2] < grid_size){
      prior[dat[i,1], dat[i,2]] <- prior[dat[i,1], dat[i,2]] + 1
    }
    
  }
  prior <- prior/sum(prior)
  return(prior)
}

generate_fisherman <- function(grid_size){
  
  # Function to generate the true location of the lost fisherman.
  # This should not effect the search decision in any way!! It is unkown
  # to the search crew.
  # Args: 
  #       grid_size: the dimensions of the square search grid

  
  mu_vec  <- c(grid_size/2, grid_size/2)
  sig_mat <- matrix(c(2, 1, 5, 5), 2,2)
  
  location  <- round(mvrnorm(1, mu_vec, sig_mat))
  true_grid <- matrix(rep(0, grid_size^2), grid_size, grid_size)
  true_grid[location[1], location[2]] <- 1
  
  return(true_grid)
}

#### Simulation ####

search_size <- 20
unifs <- runif(search_size^2, min = 0.6, max = 0.9)
detect_pr <- matrix(unifs, ncol = search_size)

#...

```

### Gibbs Sampling

-   define max iterations

-   starting values for $\beta$ and $\sigma^2$ to initialize

-   store each pair that gets sampled

```{r}
#For the inverse gamma: rinvgamma()
#install.packages("extraDistr")
library(extraDistr)
```

### Helper functions

```{r}
b_tilde = matrix(0,nrow = 5,1)

M = diag(1,5)

mu_b_function = function(M, X, b_hat, b_tilde){
  solve((M + t(X)%*%X)) %*% (t(X)%*%X%*%b_hat)
}


A2_function = function(y,b_tilde,mu_b){
  (t(y)%*%y)  - (t(mu_b)%*%(M+t(X)%*%X)%*%mu_b)
}

b_hat_function= function(X,Y){
  solve(t(X)%*%X)%*%t(X)%*%Y
}


b_hat = b_hat_function(X,Y)

b_hat

mu_b = mu_b_function(M,X,b_hat,b_tilde)

mu_b

A2 = A2_function(Y,b_tilde,mu_b)

A2
```

### Conditional Posteriors

```{r}
beta_posterior = function(mu_b,sigma){
  var = sigma*solve((M+t(X)%*%X))
}

sigma_posterior = function(a,b,A2) {
  
}
```

### Sampling from Inverse Gamma

-   Use to get $\sigma^2$

```{r}
#Fixed for all iterations
a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)


sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
```

### Sample from Multivariate Normal

-   To get $\beta$

```{r}
sigma = sigma_hat*solve((M+t(X)%*%X))
sigma
mean = mu_b
mean
beta_hat <- rmvnorm(n = 1, mean , sigma)
beta_hat

sample = c(sigma_hat,beta_hat)
names(sample) = c( "sigma", "b0", "b1", "b2", "b3", "b4")
sample
```

## Gibbs Sampling Function

```{r}
gibs_sampler = function(iterlim){
  
  #Create matrix to store values 
  sample_matrix = matrix(0,nrow = iterlim,ncol = 6)
  colnames(sample_matrix) = c("sigma", "b0", "b1", "b2", "b3", "b4")
  
  #Initialise values of sigma^2 and beta 
  a = 1
  b = 1
  shape = a + (n/2)
  rate = b + (A2/2)
  sigma_hat <- rinvgamma(1, shape, rate)
  
  sigma = sigma_hat*solve((M+t(X)%*%X))
  mean = mu_b
  beta_hat = rmvnorm(n = 1, mean , sigma)
  
  #Store intial values
  sample = c(sigma_hat,beta_hat)
  
  sample_matrix[1,] = sample
  
  
  
  for(i in 2:iterlim){
    
    #Generate value for sigma^2
    sigma_hat <- rinvgamma(1, shape, rate)
    
    #Generate value for beta 
    sigma = sigma_hat*solve((M+t(X)%*%X))
    mean = mu_b
    beta_hat = rmvnorm(n = 1, mean , sigma)
    
    #Store sample
    sample = c(sigma_hat,beta_hat)
  
    sample_matrix[i,] = sample
    
    
  }
  
  return(sample_matrix)
    
}
```

## Generate Sample

```{r}
mat = gibs_sampler(50000)

mat
```

## Trace plots

-   Need to extract all the seperate columns

    -   sigma\^2

    -   b0

    -   b1

    -   b2

    -   b3

    -   b4

### $\sigma^2$

```{r}
x = seq(1,50000)
sigma_squared = mat[,1]

plot(x,sigma_squared,type = "l")
```

### $\beta_0$

```{r}
b0 = mat[,2]
plot(x,b0,type = "l")
```

### $\beta_1$

```{r}
b1 = mat[,3]
plot(x,b1,type = "l")
```

### $\beta_2$

```{r}
b2 = mat[,4]
plot(x,b2,type = "l")
```

### $\beta_3$

```{r}
b3 = mat[,5]
plot(x,b3,type = "l")
```

### $\beta_4$

```{r}
b4 = mat[,6]
plot(x,b4,type = "l")
```

## Density Plots

### $\sigma^2$

```{r}
hist(sigma_squared,breaks = 20)
```

### $\beta_0$

```{r}
hist(b0,breaks = 20)
```

### $\beta_1$

### $\beta_2$

### $\beta_3$

### $\beta_4$

