---
title: "Assignment Template"
format: html
---

#Loading Libraries

```{r}
# install.packages("invgamma")
# install.packages("extraDistr")

library(invgamma)
library(mvtnorm)
library(ggplot2)
library(MASS)
library(extraDistr)
```

```{r}
# Packages required
require(MASS)

# Load the Auto MPG dataset
auto <- read.csv("auto.csv")

auto <- na.omit(auto)

Y <- matrix(auto$mpg, ncol = 1)

Xpreds <- auto[, c("displacement", "horsepower", "weight", "acceleration")]

X <- as.matrix(cbind(1, scale(Xpreds)))

n <- nrow(X)

```

### Gibbs Sampling

-   define max iterations

-   starting values for $\beta$ and $\sigma^2$ to initialize

-   store each pair that gets sampled

### Helper functions

```{r}
b_tilde = matrix(0,nrow = 5,1)

M = diag(1,5)

mu_b_function = function(M, X, b_hat, b_tilde){
  solve((M + t(X)%*%X)) %*% (t(X)%*%X%*%b_hat)
}


A2_function = function(y,b_tilde,mu_b){
  (t(y)%*%y)  - (t(mu_b)%*%(M+t(X)%*%X)%*%mu_b)
}

b_hat_function = function(X,Y){
  solve(t(X)%*%X)%*%t(X)%*%Y
}


b_hat = b_hat_function(X,Y)

mu_b = mu_b_function(M,X,b_hat,b_tilde)

A2 = A2_function(Y,b_tilde,mu_b)
```

### Conditional Posteriors

```{r}
beta_posterior = function(mu_b,sigma){
  var = sigma*solve((M+t(X)%*%X))
}

sigma_posterior = function(a,b,A2) {
  
}
```

### Sampling from Inverse Gamma

-   Use to get $\sigma^2$

```{r}
#| eval: false
#| include: false
#Fixed for all iterations
a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)


sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
sigma_hat <- rinvgamma(1, shape, rate)
```

### Sample from Multivariate Normal

-   To get $\beta$

```{r}
#| eval: false
#| include: false
sigma = sigma_hat*solve((M+t(X)%*%X))
sigma
mean = mu_b
mean
beta_hat <- rmvnorm(n = 1, mean , sigma)
beta_hat

sample = c(sigma_hat,beta_hat)
names(sample) = c( "sigma", "b0", "b1", "b2", "b3", "b4")
sample
```

## Gibbs Sampling Function

```{r}
gibs_sampler = function(iterlim,A2,mu_b,M){
  
  #Create matrix to store values 
  sample_matrix = matrix(0,nrow = iterlim,ncol = 6)
  colnames(sample_matrix) = c("sigma", "b0", "b1", "b2", "b3", "b4")
  
  #Initialise values of sigma^2 and beta 
  a = 1
  b = 1
  shape = a + (n/2)
  rate = b + (A2/2)
  sigma_hat <- rinvgamma(1, shape, rate)
  
  sigma = sigma_hat*solve((M+t(X)%*%X))
  mean = mu_b
  beta_hat = rmvnorm(n = 1, mean , sigma)
  
  #Store intial values
  sample = c(sigma_hat,beta_hat)
  
  sample_matrix[1,] = sample
  
  
  
  for(i in 2:iterlim){
    
    #Generate value for sigma^2
    sigma_hat <- rinvgamma(1, shape, rate)
    
    #Generate value for beta 
    sigma = sigma_hat*solve((M+t(X)%*%X))
    mean = mu_b
    beta_hat = rmvnorm(n = 1, mean , sigma)
    
    #Store sample
    sample = c(sigma_hat,beta_hat)
  
    sample_matrix[i,] = sample
    
    
  }
  
  return(sample_matrix)
    
}
```

## Generate Sample

```{r}
mat = gibs_sampler(50000,A2,mu_b,M)
```

## Trace plots $\sigma^2$

```{r}
x = seq(1,50000)
sigma_squared = mat[,1]

#png("output/plots/trace_sigma.png")
plot(x,sigma_squared,type = "l", main = expression(paste("Trace Plot for" ~ sigma^2)),
     ylab = expression(sigma^2), 
     xlab = "Iteration")
#dev.off()
```

### $\beta_0$

```{r}
b0 = mat[,2]
plot(x,b0,type = "l", main = expression(paste("Trace Plot for" ~ beta[0])),
     ylab = expression(beta[0]), 
     xlab = "Iteration")
```

### $\beta_1$

```{r}
b1 = mat[,3]
plot(x,b1,type = "l", main = expression(paste("Trace Plot for" ~ beta[1])),
     ylab = expression(beta[1]), 
     xlab = "Iteration")
```

### $\beta_2$

```{r}
b2 = mat[,4]
plot(x,b2,type = "l", main = expression(paste("Trace Plot for" ~ beta[2])),
     ylab = expression(beta[2]), 
     xlab = "Iteration")
```

### $\beta_3$

```{r}
b3 = mat[,5]
plot(x,b3,type = "l", main = expression(paste("Trace Plot for" ~ beta[3])),
     ylab = expression(beta[3]), 
     xlab = "Iteration")
```

### $\beta_4$

```{r}
b4 = mat[,6]
plot(x,b4,type = "l", main = expression(paste("Trace Plot for" ~ beta[4])),
     ylab = expression(beta[4]), 
     xlab = "Iteration")
```

## Density Plots

### $\sigma^2$

-   Add the OLS

```{r}
hist(sigma_squared, probability = TRUE,breaks = 50, 
     main = expression(paste("Posterior of" ~ sigma^2)) , xlab = expression(sigma^2))

a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)

#Overlay Inverse Gamma
curve(dinvgamma(x, shape,rate),
      from = 20, to = 45,
      add = TRUE, col = "red", lwd = 2)

abline(v = mean(sigma_squared), col = "blue", lwd = 2, lty = 2)
```

### $\beta_0$

```{r}
hist(b0,breaks = 50, main = expression(paste("Density plot of" ~ beta[0[]])) , xlab = expression(beta[0]))

#Posterior mean
abline(v = mean(b0), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[1], col = "red", lwd = 2, lty = 2)

```

### $\beta_1$

```{r}
hist(b1,breaks = 50, main = expression(paste("Density plot of" ~ beta[1])) , xlab = expression(beta[1]))

#Posterior mean
abline(v = mean(b1), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[2], col = "red", lwd = 2, lty = 2)
```

### $\beta_2$

```{r}
hist(b2,breaks = 50, main = expression(paste("Density plot of" ~ beta[2])) , xlab = expression(beta[2]))

#Posterior mean
abline(v = mean(b2), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[3], col = "red", lwd = 2, lty = 2)
```

### $\beta_3$

```{r}
hist(b3,breaks = 50, main = expression(paste("Density plot of" ~ beta[3])) , xlab = expression(beta[3]))

#Posterior mean
abline(v = mean(b3), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[4], col = "red", lwd = 2, lty = 2)
```

### $\beta_4$

```{r}
hist(b4,breaks = 50, main = expression(paste("Density plot of" ~ beta[4])) , xlab = expression(beta[4]))

#Posterior mean
abline(v = mean(b4), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[5], col = "red", lwd = 2, lty = 2)
```

## Credibility Intervals

### $\sigma^2$

```{r}
cred_ints = quantile(mat[, "sigma"], probs = c(0.025, 0.975))
```

### $\beta$

```{r}
cred_int_B = apply(mat[, c("b0","b1","b2","b3","b4")], 2, quantile, probs = c(0.025, 0.975))
```

## Confidence Intervals

-   Fit linear model and compute confidence intervals

```{r}
fit = lm(Y~X)

summary(fit)

confints = confint(fit)
```

## Density Plots w/ mean, OLS, Cred Intervals and CI

### Overlayed Density plot of $\sigma^2$

```{r}
hist(sigma_squared, probability = TRUE,breaks = 50, 
     main = expression(paste("Posterior of" ~ sigma^2)) , xlab = expression(sigma^2))

a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)

#Overlay Inverse Gamma
curve(dinvgamma(x, shape,rate),
      from = 14, to = 40,
      add = TRUE, col = "red", lwd = 2)

abline(v = mean(sigma_squared), col = "blue", lwd = 2, lty = 2)
abline(v = cred_ints, col = "darkmagenta", lwd =2, lty = 2)
```

### Overlayed Density Plot for $\beta_0$

```{r}
hist(b0,breaks = 50, main = expression(paste("Density plot of" ~ beta[0[]])) , xlab = expression(beta[0]))

#Posterior mean
abline(v = mean(b0), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[1], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 1], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[1, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_1$

```{r}
hist(b1,breaks = 50, main = expression(paste("Density plot of" ~ beta[1])) , xlab = expression(beta[1]))

#Posterior mean
abline(v = mean(b1), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[2], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 2], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[3, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_2$

```{r}
hist(b2,breaks = 50, main = expression(paste("Density plot of" ~ beta[2])) , xlab = expression(beta[2]))

#Posterior mean
abline(v = mean(b2), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[3], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 3], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[4, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_3$

```{r}
hist(b3,breaks = 50, main = expression(paste("Density plot of" ~ beta[3])) , xlab = expression(beta[3]))

#Posterior mean
abline(v = mean(b3), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[4], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 4], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[5, ], col = "darkmagenta", lwd = 2, lty = 2)
```

### Overlayed Density Plot for $\beta_4$

```{r}
hist(b4,breaks = 50, main = expression(paste("Density plot of" ~ beta[4])) , xlab = expression(beta[4]))

#Posterior mean
abline(v = mean(b4), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[5], col = "red", lwd = 2, lty = 2)

#Credibility Intervals 
abline(v = cred_int_B[, 4], col = "pink", lwd = 2, lty = 2)

#Confidence Intervals 
abline(v = confints[5, ], col = "darkmagenta", lwd = 2, lty = 2)
```

## Informative Prior

```{r}
b_tilde = matrix(c(23,-3,-5,-5,0),nrow = 5,1)

b_tilde

M = 10*diag(1,5)

M

mu_b = mu_b_function(M,X,b_hat,b_tilde)

mu_b

A2 = A2_function(Y,b_tilde,mu_b)

```

### Sample

```{r}
mat2 = gibs_sampler(50000,A2,mu_b,M)

#Extract vectors 
sigma_squared = mat2[,1]
b0 = mat2[,2]
b1 = mat2[,3]
b2 = mat2[,4]
b3 = mat2[,5]
b4 = mat2[,6]
```

## Density Plots

### $\sigma^2$

-   Add the OLS

-   Add the posterior mean

```{r}
hist(sigma_squared, probability = TRUE,breaks = 50, 
     main = "Posterior of sigma^2", xlab = "sigma^2")

a = 1
b = 1
shape = a + (n/2)
rate = b + (A2/2)

#Overlay Inverse Gamma
curve(dinvgamma(x, shape,rate),
      from = 25, to = 40,
      add = TRUE, col = "red", lwd = 2)



abline(v = mean(sigma_squared), col = "blue", lwd = 2, lty = 2)
```

### $\beta_0$

```{r}
hist(b0,breaks = 50)

#Posterior mean
abline(v = mean(b0), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[1], col = "red", lwd = 2, lty = 2)

```

### $\beta_1$

```{r}
hist(b1,breaks = 50)

#Posterior mean
abline(v = mean(b1), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[2], col = "red", lwd = 2, lty = 2)
```

### $\beta_2$

```{r}
hist(b2,breaks = 50)

#Posterior mean
abline(v = mean(b2), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[3], col = "red", lwd = 2, lty = 2)
```

### $\beta_3$

```{r}
hist(b3,breaks = 50)

#Posterior mean
abline(v = mean(b3), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[4], col = "red", lwd = 2, lty = 2)
```

### $\beta_4$

```{r}
hist(b4,breaks = 50)

#Posterior mean
abline(v = mean(b4), col = "blue", lwd = 2, lty = 2)

#OLS Estimate
abline(v = b_hat[5], col = "red", lwd = 2, lty = 2)
```

# Question 2

## Data Generating Function

```{r}
#### Data Generating Functions ####


generate_lost <- function(grid_size, nsims){
  
  # Function to generate the prior distribution for the 
  # location of the lost fisherman
  # Args: 
  #       grid_size: the dimensions of the square search grid
  #       nsims: number of samples to base the prior distribution on
  
  mu_vec  <- c(grid_size/2, grid_size/2)
  sig_mat <- matrix(c(2, 1, 5, 5), 2,2)
  
  dat <- mvrnorm(nsims, mu_vec, sig_mat)
  dat <- round(abs(dat))
  
  prior <- matrix(rep(0,grid_size^2), grid_size, grid_size)
  for (i in 1:NROW(dat)){
    
    if (dat[i,1] < grid_size & dat[i,2] < grid_size){
      prior[dat[i,1], dat[i,2]] <- prior[dat[i,1], dat[i,2]] + 1
    }
    
  }
  prior <- prior/sum(prior)
  return(prior)
}


generate_fisherman <- function(grid_size){
  
  # Function to generate the true location of the lost fisherman.
  # This should not effect the search decision in any way!! It is unkown
  # to the search crew.
  # Args: 
  #       grid_size: the dimensions of the square search grid

  
  mu_vec  <- c(grid_size/2, grid_size/2)
  sig_mat <- matrix(c(2, 1, 5, 5), 2,2)
  
  location  <- round(mvrnorm(1, mu_vec, sig_mat))
  true_grid <- matrix(rep(0, grid_size^2), grid_size, grid_size)
  true_grid[location[1], location[2]] <- 1
  
  return(true_grid)
}

#### Simulation ####

search_size <- 20
unifs <- runif(search_size^2, min = 0.6, max = 0.9)

#pi
detect_pr  <- matrix(unifs, ncol = search_size)



```

## Simulation

1.  Gain prior distribution for the fisherman's location from experts (generate lost)
2.  Gain a detection distribution given the location (basically given i)
3.  Merge
4.  Perform search of the with highest detection probability
    1.  Select grid with highest detection probability
5.  If fisherman is not detected, update the probability distribution via bayes theorem (basically update detection_probability)
    1.  Use equation 2 and 3
        1.  If statement required
6.  Continue until fisherman is found
    1.  While loop

### Generate fisherman's true location

```{r}
#Store in here
true_loc = generate_fisherman(20)
```

### Gain prior distribution for fishermen's location using expert

```{r}
expert_prior = generate_lost(20,100)
```

### Detection probability

```{r}
str(detect_pr)
```

### Combining distributions

-   Shows probabilities of successful detections

```{r}
merged_dists = expert_prior*detect_pr

```

## Search process

-   Search cell with highest probability of successful detection

### Loop

```{r}
expert_prior = generate_lost(20,100)

result = 0

numiterations = 0

while(result == 0){
  #Update the number of iterations
  numiterations = numiterations + 1
  
  #Merge distributions 
  merged_dists = expert_prior*detect_pr

  #Search cell with highest probability of occurence
  
  #Coordinates of that cell 
  highest_prob_coord = arrayInd(which.max(merged_dists),dim(merged_dists))


  #Detection probability
  pi = detect_pr[highest_prob_coord]
  
  #True state
  yi = true_loc[highest_prob_coord]
  
  #Simulate search process using bernoulli distribution
  #Zi|Yi
  result = rbinom(1,1,yi*pi)
  
  #Update occurence probabilities if fishermen not found
  
  #Will technically exit loop as soon as result = 1
  if(result == 0){
  old_theta = expert_prior[highest_prob_coord]
  
      
  # Update all other cells first (equation 3)
  expert_prior = expert_prior / (1 - pi * old_theta)
  
    
  #Update current cell (equation 2)
  expert_prior[highest_prob_coord] = (1 - pi) * old_theta / (1 - pi * old_theta)
  }
    
  
}

result
numiterations

```
